Deciphering Key Determinants of Biden Voter Support: Insights from Logistic Regression Analysis in the 2020 US Presidential Election

### 2024-03-17 paper/paper.qmd
test
added stats table

---
title: "Deciphering Key Determinants of Biden Voter Support: Insights from Logistic Regression Analysis in the 2020 US Presidential Election"
author: 
  - Fares Alkorani
  - Owais Zahid
thanks: "Code and data are available at: https://github.com/FFFiend/linear_model_investigation."
date: today
date-format: long
abstract: "In this paper, we discuss the correlation between various statistics of the respondents to the CES2020 election survey, and whether they voted for Trump or Biden. The statistics are: age, education, gender and race."
format: pdf
number-sections: true
bibliography: references.bib
---
# Introduction

@@

```{r}
#| echo: false
#| label: raw_data
#| tbl-cap: A sample of CES 2020 survey data.

@@

# formatting data, mapping race ids to races.
ces2020 <-
  ces2020 |>
  filter(votereg == 1,
         CC20_410 %in% c(1, 2)) |>
  mutate(
    voted_for = if_else(CC20_410 == 1, "Biden", "Trump"),
    voted_for = as_factor(voted_for),
    gender = if_else(gender == 1, "Male", "Female"),
    education = case_when(
      educ == 1 ~ "No HS",
      educ == 2 ~ "High school graduate",
      educ == 3 ~ "Some college",
      educ == 4 ~ "2-year",
      educ == 5 ~ "4-year",
      educ == 6 ~ "Post-grad"
    ),
    education = factor(
      education,
      levels = c(
        "No HS",
        "High school graduate",
        "Some college",
        "2-year",
        "4-year",
        "Post-grad"
      )
    ),
    race = case_when(
      race == 1 ~ "White",
      race == 2 ~ "Black",
      race == 3 ~ "Hispanic",
      race == 4 ~ "Asian",
      race == 5 ~ "Native American",
      race == 6 ~ "Middle Eastern",
      race == 7 ~ "Two or more races",
      race == 8 ~ "Other"
    ),
    race = factor(
      race,
      levels = c(
        "White",
        "Black",
        "Hispanic",
        "Asian",
        "Native American",
        "Middle Eastern",
        "Two or more races",
        "Other"
      )
    )
  ) |> select(voted_for, gender, education, race, age)
# preview data after mutation.
kable(head(ces2020))
```
[CROSS REFERENCE: FIGURE 1] shows that

>>>>

---
title: "Deciphering Key Determinants of Biden Voter Support: Insights from Logistic Regression Analysis in the 2020 US Presidential Election"
author: 
  - Fares Alkorani
  - Owais Zahid
thanks: "Code and data are available at: https://github.com/FFFiend/linear_model_investigation."
date: today
date-format: long
abstract: "On November 4, 2020, the 59th United States presidental election took place where Joe Biden became elected as 46th president of the United States. Using responses from the CES 2020 election survey (n = 8000), we study the effect that gender, race, age, and education have on whether an American voted for Joe Biden using a logistic regression model. We find that men are less likely to vote for Biden than women, respondents with higher education level are more likely to vote for Biden, and that Black respondents are most likely to vote for Biden from all other racial groups. We also find a weak negative correlation between a respondents age and whether they vote for Biden. As the upcoming United States presidental election is anticipated to be between Donald J. Trump and Joe Biden, these results have implications for how both candidates will use voter demographics to guide their campaigns."
format: pdf
number-sections: true
bibliography: references.bib
---
# Introduction

@@

```{r}
#| warning: false 
#| echo: false
#| label: raw_data
#| tbl-cap: A sample of CES 2020 survey data.

@@

# formatting data, mapping race ids to races.
ces2020 <-
  ces2020 |>
  filter(votereg == 1,
         CC20_410 %in% c(1, 2)) |>
  mutate(
    voted_for = if_else(CC20_410 == 1, "Biden", "Trump"),
    voted_for = as_factor(voted_for),
    gender = if_else(gender == 1, "Male", "Female"),
    education = case_when(
      educ == 1 ~ "No HS",
      educ == 2 ~ "High school graduate",
      educ == 3 ~ "Some college",
      educ == 4 ~ "2-year",
      educ == 5 ~ "4-year",
      educ == 6 ~ "Post-grad"
    ),
    education = factor(
      education,
      levels = c(
        "No HS",
        "High school graduate",
        "Some college",
        "2-year",
        "4-year",
        "Post-grad"
      )
    ),
    race = case_when(
      race == 1 ~ "White",
      race == 2 ~ "Black",
      race == 3 ~ "Hispanic",
      race == 4 ~ "Asian",
      race == 5 ~ "Native American",
      race == 6 ~ "Middle Eastern",
      race == 7 ~ "Two or more races",
      race == 8 ~ "Other"
    ),
    race = factor(
      race,
      levels = c(
        "White",
        "Black",
        "Hispanic",
        "Asian",
        "Native American",
        "Middle Eastern",
        "Two or more races",
        "Other"
      )
    )
  ) |> select(voted_for, gender, education, race, age)
# preview data after mutation.
kable(head(ces2020))
```
[CROSS REFERENCE: FIGURE 1] shows that
```{r}
#| warning: false
#| echo: false
#| label: dist
#| tbl-cap: The distribution of presidential preferences, by race and education.
ces2020 |>
  ggplot(aes(x = education, fill = voted_for)) +
  stat_count(position = "dodge") +
  facet_wrap(facets = vars(race)) +
  theme_minimal() +
  labs(
    x = "Highest education",
    y = "Number of respondents",
    fill = "Voted for"
  ) +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```
